% Minimal pandoc template for JCAP (jcappub.sty on article class).
% Engine: xelatex (STIX Two Text) so the prose Unicode + math render.
\documentclass[11pt,a4paper]{article}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else
  \usepackage{fontspec}
  \setmainfont{STIX Two Text}
\fi

\usepackage{jcappub}   % loads amsmath, amssymb, graphicx, natbib[numbers], hyperref
\usepackage{orcidlink} % ORCID icon next to the author (hyperref already loaded by jcappub)

% --- prose Unicode operators -> math (works under xe/pdf latex) ---
\usepackage{newunicodechar}
\newunicodechar{→}{\ensuremath{\rightarrow}}
\newunicodechar{∂}{\ensuremath{\partial}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{∎}{\ensuremath{\blacksquare}}
\newunicodechar{□}{\ensuremath{\Box}}
\newunicodechar{∝}{\ensuremath{\propto}}
\newunicodechar{∞}{\ensuremath{\infty}}
\newunicodechar{∫}{\ensuremath{\int}}
\newunicodechar{≈}{\ensuremath{\approx}}
\newunicodechar{≠}{\ensuremath{\neq}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{≤}{\ensuremath{\leq}}
\newunicodechar{≥}{\ensuremath{\geq}}
\newunicodechar{≪}{\ensuremath{\ll}}
\newunicodechar{≫}{\ensuremath{\gg}}
\newunicodechar{≲}{\ensuremath{\lesssim}}
\newunicodechar{⊙}{\ensuremath{\odot}}
\newunicodechar{⊗}{\ensuremath{\otimes}}
\newunicodechar{⟹}{\ensuremath{\Longrightarrow}}
\newunicodechar{⟺}{\ensuremath{\Longleftrightarrow}}
\newunicodechar{⇒}{\ensuremath{\Rightarrow}}
\newunicodechar{⇔}{\ensuremath{\Leftrightarrow}}
\newunicodechar{𝓗}{\ensuremath{\mathcal{H}}}
\newunicodechar{√}{\ensuremath{\surd}}
\newunicodechar{∼}{\ensuremath{\sim}}
\newunicodechar{∇}{\ensuremath{\nabla}}
\newunicodechar{↔}{\ensuremath{\leftrightarrow}}
\newunicodechar{⟨}{\ensuremath{\langle}}
\newunicodechar{⟩}{\ensuremath{\rangle}}
\newunicodechar{−}{\ensuremath{-}}
\newunicodechar{ℓ}{\ensuremath{\ell}}
\newunicodechar{ħ}{\ensuremath{\hbar}}
\newunicodechar{′}{\ensuremath{{}^\prime}}
\newunicodechar{Ḣ}{\ensuremath{\dot H}}
\newunicodechar{Ṡ}{\ensuremath{\dot S}}
\newunicodechar{₊}{\ensuremath{{}_{+}}}

% --- pandoc body requirements ---
\usepackage{longtable,booktabs,array}
\usepackage{calc}
\usepackage{float}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{%
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi%
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
\makeatother
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\graphicspath{{../../results/paper/}{../../results/}}

% --- front matter (jcappub macros) ---
\title{Supplementary Material for ``Structural Entropy Dark Energy: a
fixed-parameter, growth-gated holographic dark-energy model without a
cosmological constant''}
\author{Stilian Pandev\,\orcidlink{0009-0005-8153-071X}}
\affiliation{Independent researcher}
\emailAdd{stilianpandev@gmail.com}
\abstract{This supplement collects the supporting motivations and
numerical probes for the volume-law horizon postulate, demoted from the
main article to keep its main line focused: three further consistency
motivations for the volume endpoint, the Majorana--SYK diagonalisation
that fixes the horizon state, numerical probes of the state and counting
halves of the postulate, the degrees-of-freedom accounting that locates
the postulate as a counting (not thermalisation) claim --- with the SYK
and causal-set comparisons --- and the summary of how far the postulate
is derivable. None is a derivation; each is a consistency check.
Cross-references of the form §N refer to sections of the main article;
§S-N to sections of this supplement.}


% absorb small unbreakable-line overruns instead of margin overflow
\setlength{\emergencystretch}{3em}

\begin{document}
\maketitle
\section{Supporting motivations and numerical probes for the volume-law
postulate}\label{sec-1}

This supplement collects the material demoted from §8 of the main
article to keep the main line focused: three further consistency
\emph{motivations} for the volume endpoint (§8.1) and the
\emph{numerical probes} of the state and counting halves of the
postulate (§S2). None is a derivation; each is a consistency check that
makes the postulate less \emph{ad hoc} without selecting it. (The CHR
near-critical layer, the cross-horizon black-hole reading, and the
Verlinde dark-matter connection are developed in a separate foundations
companion paper \citep{Pandev:2026foundations}.)

\subsection{Three further consistency motivations for the volume
endpoint}\label{sec-1-1}

\emph{(supplementing the geometric ceiling, the CKN scale/form split,
and the direct area-law disfavouring of §8.1)}

\begin{itemize}
\tightlist
\item
  \textbf{A driven steady state.} The horizon need not self-thermalise
  --- its scrambling time (∼ ln \(S / H \approx  280/H\)) far exceeds
  its age. Instead the structure-growth gate (the rise of
  \(f_{\mathrm{sat}}\)) \emph{drives} it off the ground state; with a
  scrambling-limited relaxation rate the volume-law state is
  \emph{maintained} as a non-equilibrium steady state throughout
  structure formation (the long scrambling time becomes an asset). The
  volume law is then a fixed point of SEDE's own dynamics, not an
  external input.
\item
  \emph{A roughening universality class} (\textbf{withdrawn}). This
  motivation modelled the horizon as a stochastic growing surface,
  \(\partial h/\partial t = \nu \nabla ^{2}h + (\lambda _{\mathrm{K}}/2)(\nabla h)^{2} + \eta\),
  and argued the deformation was fixed by the roughening universality
  class. It is retracted: the count companion paper
  \citep{Pandev:2026count} shows the classical apparent horizon is
  constraint-slaved --- \(\theta\)₊ = 0 is elliptic, with no growth
  equation at any order (its eq. 2.9) --- and has no minimum-cut
  surface, so it does not roughen and there is no height field or
  roughening universality class. The ceiling \(\Delta  \leq  1\) and the
  value \(\Delta  = 1\) rest instead on the count paper's leg-budget
  capacity theorem and on birth-selection (foundations §5), not on any
  surface-growth picture.
\item
  An independent emergent-gravity proposal \citep{Verlinde:2016toy}. In
  Verlinde's programme the de Sitter vacuum carries a volume-law
  contribution to its entanglement entropy,
  \(S \propto  V_{\mathrm{H}}\), in addition to the horizon area law ---
  the same super-area scaling SEDE adopts at \(\Delta  = 1\). A second,
  independently-motivated route arriving at a volume-law entropy makes
  the postulate less \emph{ad hoc}; it remains a motivation, not a
  proof, since Verlinde's volume-law derivation is itself heuristic.
\end{itemize}

\subsection{The state is maximally entangled (Majorana-SYK
diagonalisation)}\label{sec-1-2}

The \emph{state} half of the postulate (§S2) --- that the horizon dof
are maximally entangled (thermal), not in a low-entanglement ground
state --- is settled and not peculiar to SEDE. Exact diagonalisation of
the Majorana Sachdev--Ye--Kitaev model
\citep{Sachdev:1992fk, Kitaev:2015}, the canonical chaotic,
black-hole-dual system saturating the Maldacena--Shenker--Stanford bound
\(\lambda _{\mathrm{L}} \leq  2\pi T\) \citep{Maldacena:2015waa} (V60),
confirms: the gap ratio reproduces Wigner--Dyson statistics with the
correct N-mod-8 symmetry class
\citep{GarciaGarcia:2016mno, Cotler:2016fpe, Atas:2012np}, eigenstates
saturate the Page value \citep{Page:1993df, Vidmar:2017pak}, and the
OTOC scrambles completely --- while an integrable control does none of
these. But this \textbf{does not fix \(\Delta\)}: a black hole is a
\emph{maximal} scrambler yet area-law (S = A/4, reproduced by string
microstate counts), and SYK, being all-to-all (geometry-free), cannot
even pose the counting question.

\emph{(The SYK scrambling figure --- the gap ratio climbing through the
N-mod-8 RMT classes, the Page-value saturation of the eigenstate
entanglement, and the complete OTOC scrambling of the q=4 model against
the q=2 integrable control --- is omitted from this version; the
underlying diagnostics are the verification-suite check V60 and
regenerate from the accompanying code.)}

\subsection{The counting question --- numerical probes (allowed,
realisable, not selected)}\label{sec-1-3}

The \emph{counting} half (§S2) --- does the horizon dof number grow as
area \((N \propto  R^{d-2}\), \(\Delta =0\)) or bulk volume
\((N \propto  R^{d-1}\), \(\Delta =1\))? --- is where SEDE's postulate
lives. Three numerical probes (V62, V61, V63):

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
question
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
probe
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
result
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Allowed? by the established entropy bounds & volume entropy vs the
smooth-area bounds & \(S \propto  R^{3}\) overshoots the
holographic/Bousso/Bekenstein bound A/4 by \(\propto  R\) (∼\(10^{61}\)
at the cosmic horizon); consistent only as a genuine
\(d_{\mathrm{H}}=3\) fractal horizon (true area \(\propto  R^{3}\) ⟹
\(S = A_{\mathrm{fractal}}/4\) \emph{exactly} saturates the bound) \\
Selected? by quantum gravity & causal-set sprinkle, Minkowski and de
Sitter & both scalings present; the canonical horizon object --- causal
\emph{links} --- is area \((R^{d-2}\), the Bekenstein--Hawking
recovery), in flat space and, because dS is conformally flat,
identically in de Sitter. \textbf{Volume is not the default.} (measured:
bulk count \(1.92\approx 2\) ⟹ \(\Delta =1\); link count
\(0.98\approx 1\) ⟹ \(\Delta =0\), in 2+1D) \\
Realisable? in a holographic code & Ryu--Takayanagi min-cut, local vs
nonlocal & local (lattice) connectivity ⟹ area law (exponent 1.0);
nonlocal (expander) connectivity ⟹ volume law (exponent 1.9).
\(\Delta =1\) is realisable; its required ingredient is \emph{nonlocal}
horizon connectivity (the all-to-all structure of SYK). \\
\end{longtable}

The pattern: volume-counting is allowed (not by evading the bounds but
by the horizon being genuinely space-filling --- the postulate itself),
realisable (a nonlocal holographic code produces \(S \propto  R^{d-1}\)
exactly), but not selected (the frameworks we can compute in default to
area for the canonical horizon object, and a black hole is area-law). No
consistency argument available to us \emph{derives} the volume count;
equally, none \emph{excludes} it. The two deciders are stated in §S2:
the dS static-patch Hilbert-space dimension (theoretical) and the
deformation \(\Delta\) (empirical, §5.6/§6).

\section{\texorpdfstring{The postulate is a dof-\emph{counting} claim,
not a thermalisation
claim}{The postulate is a dof-counting claim, not a thermalisation claim}}\label{sec-2}

The phrase ``volume-law'' conflates two claims of very different status.
\textbf{(i) The state:} are the horizon degrees of freedom maximally
entangled (thermal) or in a low-entanglement ground state? \textbf{(ii)
The counting:} does the \emph{number} of horizon dof grow as the area,
\(N \propto  R^{d-2}\), or the spatial volume, \(N \propto  R^{d-1}\)?
Since S ∼ (state factor) × N, the deformation \(\Delta\) --- through
\(S \propto  A^{1+\Delta /2}\)∝ \(R^{(d-2)(1+\Delta /2)}\) --- is fixed
by the counting (area→\(\Delta =0\), volume→\(\Delta =1\) in d=4); the
state only sets whether the prefactor is maximal. SEDE's postulate is,
precisely, the counting claim.

\emph{The state half is settled, and is not peculiar to SEDE.} Maximal
scrambling has a precise meaning in the Sachdev--Ye--Kitaev model
\citep{Sachdev:1992fk, Kitaev:2015}, the canonical chaotic,
black-hole-dual system. But maximal scrambling \textbf{does not fix
\(\Delta\)}: a black hole is a \emph{maximal} scrambler yet area-law (S
= A/4, \(\Delta =0\), reproduced by string microstate counts).
Scrambling is therefore shared with the \(\Delta =0\) black hole; and
SYK, being all-to-all (geometry-free), cannot even pose the counting
question. (An explicit Majorana-SYK diagonalisation confirming the state
is maximally entangled --- Wigner--Dyson statistics, Page-value
saturation, complete OTOC scrambling --- is given in §S1.2; it settles
the \emph{state}, not the counting.)

\emph{The counting half is where SEDE lives, and area is the default.}
In causal-set quantum gravity --- the setting where the counting
question is natural --- both scalings appear, but the \emph{canonical}
horizon-entropy object, the causal-link count that recovers the
Bekenstein--Hawking law (Sorkin and collaborators), is the area one
\((R^{d-2})\), in flat space and identically in de Sitter. So even in
the framework most hospitable to volume-counting, area-law is the
default, and SEDE's \(\Delta =1\) is the specific, non-default
identification of horizon entropy with the bulk count. (The causal-set
sprinkle measurements, the entropy-bound check, and a Ryu--Takayanagi
holographic-code realisation --- volume-law requires \emph{nonlocal}
horizon connectivity --- are in §S1.3; their verdict is that the volume
count is \textbf{allowed and realisable but not selected}: no
consistency argument available to us derives it, and none excludes it.)

The sharpened postulate, properly located. The one open item is
therefore \emph{not} about thermalisation or chaos (settled, and shared
with area-law black holes) but is purely a dof-counting statement:

\begin{quote}
the cosmic horizon's entropy counts its bulk (volume-scaling,
\(N \propto  R^{3}\) ⟺ \(d_{\mathrm{H}} = 3\)) degrees of freedom rather
than its boundary (area-scaling, links) ones --- a count that exceeds
the holographic area bound by \(R/\ell _{\mathrm{P}}\) (the seam of
§8.3).
\end{quote}

It reduces to two deciders. \emph{Theoretically} it is precisely
\emph{``is the Hilbert-space dimension of the de Sitter static patch
\(e^{Area/4}\) or \(e^{Volume}\)?''} --- a frontier problem a complete
dS holography would settle. \emph{Empirically} it is the deformation
\(\Delta\): \(\Delta =0\) (area) is already disfavoured by present data
(§5.6), and DESI DR3 + Euclid will measure \(\Delta\) to ∼0.09 (§6).
SEDE's postulate is thus not an arbitrary assumption but a
\emph{consistent, realisable bet on a well-posed open question of
quantum gravity}, with a decisive test imminent. Tellingly, the one
horizon we can check today --- the black hole --- is area-law; on the
evidence here this asymmetry is a genuine conceptual cost. The count
companion paper \citep{Pandev:2026count} \emph{offers} a resolution ---
a state-dependent count in which a black hole is area-law because
\emph{undriven} and only the structure-driven cosmic horizon activates
the volume law --- which, if its (conjectural) mechanism holds, would
turn the cost into a predicted feature, with the black-hole side
derivable and a sharp falsifier (an isolated volume-law black hole).

\section{Is the postulate derivable? Summary of the reduction (full
treatment in the foundations companion paper)}\label{sec-3}

The volume-law postulate is not monolithic: it splits into a
\emph{state} (the horizon dof are maximally entangled/thermal), a
\emph{form} (the entropy scales as the volume), a \emph{scale} (the
magnitude is \(\rho _{\mathrm{crit}}\), not
\(\rho _{\mathrm{Planck}}\)), and a \emph{count} (the number of dof
grows as the bulk, \(N \propto  R^{3}\), not the boundary,
\(N \propto  R^{2}\)). The deformation \(\Delta\) is fixed by the count
alone. A systematic study of how derivable each piece is --- five routes
(de Sitter holography/SYK, entanglement first law in a thermal state,
Verlinde emergent gravity, gravitational non-additivity, and a
Bekenstein no-go), the reduction of the count to a \emph{driven
non-equilibrium steady state}, and three closure routes --- is developed
in the foundations companion paper \citep{Pandev:2026foundations}; we
state the conclusions here, as they bear on how \emph{ad hoc} the
postulate is.

The \emph{state} is first-principles (a maximal scrambler is
volume-law-entangled; §S1.2, §S2); the \emph{form} reduces to
thermalisation (volume-law entanglement is the generic behaviour of a
thermal reference state for regions larger than the thermal length); the
\emph{scale} is CKN (§8.2). What remains irreducible is purely the
count, and a Bekenstein no-go makes precise \emph{why}:
maximum-entropy/energy bounds give the \emph{area} law (§8.3), so the
volume count cannot be obtained from any energy argument --- it is a
genuine state/counting statement. The count is, in turn, downstream of
\emph{connectivity} (locally-connected horizon dof give area-law
entropy, nonlocal give volume-law) and connectivity of \emph{interaction
range}: gravity is strongly long-range (1/r with
\(\alpha  = 1 \leq  d\)), hence non-additive --- the volume class is the
\emph{natural} one for a self-gravitating horizon, with area-law arising
only when the holographic bound intervenes for an \emph{isolated,
equilibrium} horizon (a black hole). The cosmic horizon differs in being
continuously \emph{driven} by structure formation, which converts the
black-hole/cosmic asymmetry (§7) into a discriminator (equilibrium →
area, driven → volume) and reduces the residue to a single, falsifiable
\textbf{driven-NESS statement}: a strongly-long-range horizon, driven by
structure formation, settles in the volume-law steady state rather than
relaxing to the area-law equilibrium. The foundations companion paper
supplies the two mechanisms this requires --- \emph{selection at the
horizon's birth} (the N-dependent cooperative coupling
\(J/J_{\mathrm{c}} = 2\pi Gm^{2}N\) carries the horizon, grown from
Planckian size, through the ordering bifurcation, with the Barrow tilt
selecting the volume branch and long-range hysteresis locking it; no
barrier is ever crossed) and the structure \emph{drive}, whose envelope
is the gate \(f_{\mathrm{sat}}\), which \emph{populates} the selected
volume capacity rather than selecting the branch --- so the static
\emph{counting} postulate is replaced by the weaker, more physical
statement of a birth-selected branch plus derived occupancy kinetics.

Closure --- deriving even that bistable free energy from first
principles --- would reduce the dark sector to \(\Omega _{\mathrm{m}}\)
plus established physics. It takes three precise forms: a de
Sitter-holography computation of the bulk count (open, and provably not
shortcut-able through energy bounds); exhibiting the volume branch as a
real saddle/state (the bistable landscape follows from gravitational
cooperativity given that branch); or \emph{measuring} \(\Delta\). The
last is decisive and already in hand: DESI DR3 + Euclid pin \(\Delta\)
to ∼0.09 (§6), so \(\Delta  = 1\) establishes the volume count as fact
and \(\Delta  = 0\) refutes it. The postulate is thus a single,
sharply-posed, falsifiable statement with an imminent test --- not an
unconstrained assumption.

\section*{Code and data availability}\label{code-and-data-availability}
\addcontentsline{toc}{section}{Code and data availability}

All code and analysis pipelines are publicly available at
\url{https://github.com/spsingularity/sede-cosmology}; the numerical
diagnostics above are regenerated by the accompanying code. A tagged
release is archived at Zenodo, DOI
\href{https://doi.org/10.5281/zenodo.21050314}{10.5281/zenodo.21050314}.
\bibliographystyle{JHEP}
\bibliography{refs}
\end{document}
